---
title: "p8105_hw3_ha2546"
author: "Hana Akbarnejad"
date: "10/9/2019"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(viridis)
library(ggridges)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = "90%"
  )

library(p8105.datasets)
data("instacart")
```

```{r random_sample}

# set.seed(5)
# random_sample = instacart %>% 
#   filter(
#   sample(pull(instacart, order_id))
#   )

```

In the first problem, we are interested in exploring the _instacart_ dataset. This dataset stores information about online grocery shopping from 2017. The datset contains `r nrow(instacart)` observations of `r n_distinct(pull(instacart, order_id))` unique orders of `r n_distinct(pull(instacart, product_name))` differennt products. There are `r ncol(instacart)` variables associated with these rows such as _order_dow_ (the day of the week that the order has been placed), _order_hour_of_day_ (the time of the day that the order has been placed), _product_name_ (the item purchased), _aisle_ (the aisle that the item has chosen from), etc. We can take differenvt samples from this dataset.

# Problem1

How many aisles are there, and which aisles are the most items ordered from?
I grouped the dtaset by the aisles, and understood that there are total of 134 different isles. Then I counted the number of times items have been ordered from each isle. By further arranging this column in descending order, I found the most popular isles. Top five aisles with the most items ordered from are fresh vegetables, fresh fruits, packaged vegetable fruits, yogurt, packaged cheese.I then made a plot that shows the number of items ordered in each aisle (only for aisles with more than 10000 items ordered). For this purpose I made a bar plot that shows the aisle names on x axis and the number of items ordered on y axis.

**should we arrange the columns from most to least? how?**

```{r instacart}

aisle_fav_n = instacart %>% 
  group_by(aisle) %>% 
  count(aisle, name = "aisle_order_n") %>% 
  arrange(desc(aisle_order_n))

aisle_fav_plot = aisle_fav_n %>% 
  filter(aisle_order_n > 10000) %>% 
  ggplot(aes(x = reorder(aisle, -aisle_order_n), y = aisle_order_n)) +
  geom_bar(stat="identity", width=0.5, color="blue", fill="white") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(
    title = "Number of Orders in Favorite Aisles",
    x = "Aisle name",
    y = "Number of orders",
    caption = "Plot 1"
  )

aisle_fav_n
aisle_fav_plot
```

**top_n doesn't seem working**

```{r}
x = instacart %>% 
  select(product_name, aisle) %>% 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(product_name, aisle) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count)) %>% 
  top_n(3, count)
```

In this part,

```{r}
hours_days_products = instacart %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(
    hours_mean = mean(order_hour_of_day)
  ) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = hours_mean
  ) %>% 
  knitr::kable(digits = 1)
```

